---
title: "SQL Exercise"
format: html
editor: visual
---

```{r,include = FALSE}
library(DBI)
library(RSQLite)
```

```{r,include = FALSE}
con_chocolate <- DBI::dbConnect(drv = RSQLite::SQLite(),
                                 dbname = "chocolate.sqlite")
DBI::dbListTables(con_chocolate)
```

## Q1

1.  the query meaning is - create for me a table with two columns the first is a list of the sales rep’s full name, and the second is of the sum of the orders this rep did,the column will be named total_candy_bars_sold. filter those column so it will show me only rep and sum orders from 2022 and so it will show me only sales rep that joined the company at 2010. the sum function mentcend before sum by groups of salers name.

    the info will come from two tables : salesreps and orders, combined in an inner join were the column srid indicate to them what column to use to combie them.

2.  

```{sql connection = con_chocolate, output.var="q1"}
SELECT salesreps.Name, SUM(orders.amount) AS total_candy_bars_sold
FROM salesreps INNER JOIN orders USING (srid)
WHERE
orders.sale_date BETWEEN '2022-01-01' AND '2022-12-31'
AND salesreps.year_joined = 2010
GROUP BY salesreps.Name;
```

## Q2

```{sql, connection = con_chocolate, output.var = "q2"}
WITH for_now AS (
    SELECT cid, SUM(orders.amount) AS total_orders
    FROM orders
    GROUP BY cid
)
SELECT total_orders, COUNT(cid) AS N
FROM for_now
GROUP BY total_orders;
```

## Q3

the candybar with the most orders at the firstquarter of 2022 is **Cocount Crave** and **Hess, Christope**r sold the most of it in the second quarter of 2022

```{sql, connection = con_chocolate, output.var = "q3"}
WITH top_product AS (
  SELECT pid, SUM(amount) AS total_amount
  FROM orders
  WHERE sale_date BETWEEN '2022-01-01' AND '2022-03-31'
  GROUP BY pid
  ORDER BY total_amount DESC
  LIMIT 1
)
SELECT pid, candy_names, srid, salesreps.Name, SUM(amount) AS total_sales
FROM orders
INNER JOIN products USING (pid)
INNER JOIN salesreps USING (srid)
WHERE sale_date BETWEEN '2022-04-01' AND '2022-06-30'
  AND pid IN (SELECT pid FROM top_product)
GROUP BY srid, salesreps.Name
ORDER BY total_sales DESC;
```
